---
title: "R Notebook"
output: html_notebook
---

##Aim
Identifying outliers/ problem areas with regards to deliveries. 
Using clustering, I aim to recognise delivery locations that present the highest percentages of late deliveries in order to address/investigate these.
Ideally, the data will show higher numbers of deliveries leaning more heavily towards 0% late.


#Set Packages
```{r}
library(readr)
library(tidyverse)
library(ggplot2)
```




#Import Datasets
```{r}
dataset <- read_csv("Deliveries2.csv", col_types = cols(Pallets = col_integer()))
View(dataset)
```


#Workout Clusters
```{r}
kclustersMax <- 10
WSS <- sapply(1:kclustersMax,
              function(k){kmeans(Deliveries_2[c("AvgOfTrueOrFalse", "CountOfDepot")],k)$tot.withinss})
plot(1:kclustersMax, WSS,
     type = "b", pch = 19, frame = false,
     xlab = "Number of Clusters K",
     ylab = "total within clusters - Sum of squares")                
```


#Plot Graphs & Kmeans
```{r}
kclusters <- 4
kmeansresult <- kmeans(Deliveries_2[c("AvgOfTrueOrFalse", "CountOfDepot")],
                       centers = kclusters)
Deliveries_2$cluster <- kmeansresult$cluster
Deliveries_2$cluster <- as.factor(Deliveries_2$cluster)

graphed <- qplot(x= Deliveries_2$AvgOfTrueOrFalse, y = Deliveries_2$CountOfDepot, label= Deliveries_2$Postcode, color= Deliveries_2$cluster, ylab = "Number of Deliveries", xlab = "Percentage Fail Rate")

```



##Analysis Breakdown
The graph (Rplot) created displays many features of a geometric distribution, arising from the face the x axis revolves around the average of a true/false value. The more values there are, the harder the individual curves appear.

Clustering towards the left hand side of the graph suggests a greater percentage of on time deliveries.

The few outliers on the first cluster does however raise questions around these, which will be investigated. From cursory investigation, this is because they fall into key qualities of Southern Ireland (Which follow different rules), or are returns (not same limits on time).

I'm going to collate the second and third clusters into a new data frame, then try this again to get a better breakdow, and provide a closer look on the central values for addressing alongside Cluster 4's key outliers from this first analysis.



#Create new dataset
```{r}
Deliveries_3 <- data.frame(Deliveries_2$Consignor[Deliveries_2$cluster == 2 | Deliveries_2$cluster == 3], Deliveries_2$Postcode[Deliveries_2$cluster == 2 | Deliveries_2$cluster ==3], Deliveries_2$AvgOfTrueOrFalse[Deliveries_2$cluster == 2 | Deliveries_2$cluster == 3],
Deliveries_2$CountOfDepot[Deliveries_2$cluster == 2 | Deliveries_2$cluster == 3])
View(Deliveries_3)
names(Deliveries_3)[1] <- "Consignor"
names(Deliveries_3)[2] <- "Postcode"
names(Deliveries_3)[3] <- "AvgOfTrueOrFalse"
names(Deliveries_3)[4] <- "CountOfDepot"
```


#Workout Clusters 2
```{r}
kclustersMax2 <- 10
WSS2 <- sapply(1:kclustersMax2,
              function(k){kmeans(Deliveries_3[c("AvgOfTrueOrFalse", "CountOfDepot")],k)$tot.withinss})
plot(1:kclustersMax2, WSS,
     type = "b", pch = 19, frame = TRUE,
     xlab = "Number of Clusters K",
     ylab = "total within clusters - Sum of squares")                
```


#Plot Graphs & Kmeans attempt 2
```{r}
kclusters <- 4
kmeansresult2 <- kmeans(Deliveries_3[c("AvgOfTrueOrFalse", "CountOfDepot")],
                       centers = kclusters)
Deliveries_3$cluster <- kmeansresult2$cluster
Deliveries_3$cluster <- as.factor(Deliveries_3$cluster)

graphed2 <- qplot(x= Deliveries_3$AvgOfTrueOrFalse, y = Deliveries_3$CountOfDepot, label= Deliveries_3$Postcode, color= Deliveries_3$cluster, ylab = "Number of Deliveries", xlab = "Percentage Late Rate")

```


#Findings & Final Remarks

This graph (Rplot2)provides a much more specific view to spot outliers. Cluster 4 has, encouragingly, very few high number deliveries in the region of high late percentage, but 100% late for 2 outliers requires investigation. 

Overall, whilst the clustering has perhaps, rather dissapointingly, been purely based on number of deliveries. It has also highlighted some key points for further analysis and focus, wherein high deliveries - high late rate, can be properly addressed.

#Geometric Distribution
Here I am looking to find out the probability a certain delivery will be delivered late after a certain number of faliures.

If understood correctly, the probability drops as more deliveries occur. (Rplot03 in file)
```{r}

plot(Deliveries_3$CountOfDepot, dgeom(Deliveries_3$CountOfDepot,Deliveries_3$AvgOfTrueOrFalse), type="h", ylim=c(0,0.1), main="Geometric distribution for P = %Late Deliveries", ylab="Probability of Y", xlab="Y=Number of lates before first success")

```

Upon reflection, this is incorrect usage, but I'll leave it in for seeing my thought process. For fear of expanding beyond the brief, I will halt this for now.

##Next steps
After identification of outliers in cluster 4 (AKA High Deliveries, High Late Rate), I aim to extract them from the data before analysing them further individually. From this, I will then need to discern what reasons they fall into as to their lateness, and whether them being marked as late is down to a system error or a scheduling error. For example, a system error could be that all returns are marked as late, despite them having no time limit, or deliveries to southern ireland being marked as 24 hour, despite that service not existing, hence the large number of lates.
